\section{Configured LLMs}

After defining the PDDL problem set and the selected instance tiers, the next step was to define the model pool used to generate candidate plans. The benchmark framework was designed to support two execution backends: Leonardo HPC by CINECA and the NVIDIA API. Leonardo was used for local execution of Hugging Face models, while the NVIDIA API provided remote access to additional models without requiring every checkpoint to be installed and hosted locally.

\subsection{Execution Backends}

\subsubsection{Leonardo HPC}

Leonardo HPC was the main platform considered for local model execution. The models were loaded through the Hugging Face backend, allowing direct control over the execution environment, GPU resources, numerical precision, prompt formatting, and generation parameters.

The selection of locally executable models was supported by whichLLM, a repository used to estimate which LLMs can run on a given hardware configuration and to compare models by release recency. In this project it was used as a preliminary feasibility reference, not as an automatic selection rule: the final local set was still constrained by the available Leonardo environment, dependency compatibility, and model access.

\subsubsection{NVIDIA API}

The NVIDIA API was used as a remote inference backend. This reduced the setup effort required to test recent open-weight models and made it possible to include models that would have been less convenient to install or run locally.

The benchmark pipeline remained the same across both backends: prompts, parsing, validation, and scoring followed the same workflow, while only the model adapter changed. The NVIDIA API backend also introduced practical constraints, such as rate limits, temporary service availability, and model-specific API key requirements.

\subsection{Selection Criteria}

The models were selected according to four practical criteria:

\begin{itemize}
    \item \textbf{Hardware feasibility:} local models had to be compatible with the available Leonardo environment, using whichLLM as a practical reference for hardware requirements.
    \item \textbf{Recency:} priority was given to recent LLM families, using whichLLM as a reference to compare model release dates.
    \item \textbf{Model diversity:} the benchmark included models from different families to avoid evaluating a single architectural lineage.
    \item \textbf{Evaluation availability:} models included in the reported comparison had to be executable through the same prompt, parsing, validation, and scoring workflow.
\end{itemize}

Using both backends made the framework more flexible than a Leonardo-only setup. Leonardo provided controlled local execution, while the NVIDIA API allowed the benchmark to cover additional hosted models through the same interface.

\subsection{Configured Models}

Table~\ref{tab:implemented-llm-models} reports the model pool considered by the benchmark framework. For the Hugging Face backend, the table lists the active local models in the Leonardo registry. For the NVIDIA API backend, it lists the remote models for which scored benchmark outputs were available in the collected runs. Some additional endpoints may remain present in the registry for future experiments, but they are not treated as part of the analyzed model set unless they have corresponding scored outputs.

\begin{table}[H]
\centering
\small
\begin{tabularx}{\linewidth}{p{0.17\linewidth}p{0.16\linewidth}p{0.31\linewidth}X}
\toprule
\textbf{Backend} & \textbf{Execution mode} & \textbf{Model identifier} & \textbf{Model name} \\
\midrule
Leonardo HPC & Local & \path{hf_gemma_4_31b_it} & Gemma 4 31B IT \\
Leonardo HPC & Local  & \path{hf_phi_4} & Phi-4 \\
Leonardo HPC & Local & \path{hf_gpt_oss_120b} & GPT-OSS 120B \\
Leonardo HPC & Local  & \path{hf_qwen_3_6_27b} & Qwen 3.6 27B \\
\midrule
NVIDIA API & Remote hosted & \path{nvidia_phi_4_mini_instruct} & Phi-4 Mini Instruct \\
NVIDIA API & Remote hosted & \path{nvidia_gpt_oss_120b} & GPT-OSS 120B \\
NVIDIA API & Remote hosted & \path{nvidia_kimi_k2_6} & Kimi K2.6 \\
NVIDIA API & Remote hosted & \path{nvidia_llama_3_3_70b_instruct} & Llama 3.3 70B Instruct \\
NVIDIA API & Remote hosted & \path{nvidia_nemotron_3_nano_30b_a3b} & Nemotron 3 Nano 30B A3B \\
NVIDIA API & Remote hosted & \path{nvidia_nemotron_3_ultra_550b_a55b} & Nemotron 3 Ultra 550B A55B \\
NVIDIA API & Remote hosted & \path{nvidia_qwen_3_5_122b_a10b} & Qwen 3.5 122B A10B \\
\bottomrule
\end{tabularx}
\caption{LLM models configured or included in the benchmark model pool.}
\label{tab:implemented-llm-models}
\end{table}
