\section{Benchmark Output Parsing and Output Structure}

The benchmark execution pipeline produces traceable and independently analyzable artifacts. All generated data is stored under \texttt{Benchmark\_Framework/outputs}, which is divided according to the main stages of the evaluation: generation, parsing and scoring.

This separation is a core design choice. It makes it possible to distinguish between different failure modes: a model may generate text that contains a reasonable plan but is not cleanly parseable, or it may produce a syntactically valid action sequence that still fails during formal validation. The artifacts are persisted when a benchmark case terminates, either because a valid plan has been found or because the configured maximum number of iterations has been reached.

The \texttt{outputs} directory is organized into three main subdirectories:
\begin{itemize}
    \item \textbf{raw}: stores the direct model interaction and generation metadata.
    \item \textbf{parsed}: stores the structured action sequences extracted from the raw text and reasoning text.
    \item \textbf{scored}: stores validation outcomes, repair information and aggregate metrics for the case.
\end{itemize}

Each layer follows the same hierarchical path:
\begin{center}
\texttt{<run\_id>/<model>/<protocol>/<domain>/<difficulty>/<instance>.json}
\end{center}
This layout makes the three artifacts for the same benchmark instance directly comparable across the raw, parsed and scored layers.


\subsection{Raw}

The files in \texttt{outputs/raw/} are the closest record of the interaction with the LLM. Each artifact stores: 
\begin{itemize}
  \item prompts sent to the model
  \item model response
  \item any provider-side reasoning text when available
  \item generation metadata such as token usage, latency and notes
\end{itemize}

This layer is the baseline record of what the model produced before any interpretation by the parser or validator.

\subsubsection{Raw File Structure}

Listing~\ref{lst:raw_json_structure} shows the main fields stored in a raw artifact. Provider-specific generation fields may vary across backends, but the surrounding structure is shared.

\begin{lstlisting}[language=Python, caption={Structure of raw JSON file}, label={lst:raw_json_structure}]
    {
  "model_id": #LLM model selected,
  "task_id": #problem and file selected,
  "protocol_id": #iterative repair or direct plan,
  "task_family": #PDDL domain problem selected,
  "tier": #easy, medium or hard depending on the instance,
  "instance_id": #instance selected,
  "generation_time_seconds": #sum of the time for each iteration,
  "attempts": [
    {
      "iteration": #N. of current iteration,
      "generation_time_seconds": #time for the generation of the iteration,
      "messages": [ #used for the reasoning message and the text prompts
        {
          "role": "system" #define the role of whos receiving the message,
          "content": #Here we place our text prompt, for example:
          "You are a PDDL planning assistant.\n\nYour task is to produce a valid plan for the given PDDL domain and problem.\n\nGlobal rules:\n- Use only action schemas explicitly defined in the domain..."
        },
        {
          "role": "user",
          "content": "TASK FAMILY: block-grouping\nDIFFICULTY: easy\nINSTANCE: pfile1\n\nTASK FAMILY: block-grouping\n\nYou are solving a numeric PDDL planning problem..."
        }
      ],
      "generation": { #raw output generated by the LLM, example:
        "raw_text": "(move_block_left b1)\n(move_block_left b1)\n(move_block_left b1)\n(move_block_left b2)..."
        #following tokens feature insterted but not useful
          "prompt_tokens": null,
          "completion_tokens": null,
          "total_tokens": null
        },
        "latency_s": #time latency,
        "notes": [
          #notes if present
        ],
        "reasoning_text": #reasoning text given us by the LLM
        "We need to produce a plan moving blocks to satisfy goal: blocks b1,b2,b3,b4 must be at same location (both x and y equal). b5 must be at a different location..."
        "finish_reason": #finish reason for example stop of numb of max iter reached,
        "stop_reason": #for example error or stop,
        "reached_max_tokens": false,
        "stream_complete": #true or false,
        "stream_error": #error or null,
        "partial_output": #true or false,
        "timed_out_by_job_limit": #true or false,
        "generation_time_seconds": #time used for generation
      }
    },
  ]
}

\end{lstlisting}

\subsection{Parsed}

The \texttt{outputs/parsed/} directory contains the result of the first processing step after generation. The parser attempts to extract executable PDDL actions from the raw model text and stores the extracted plan in a structured form.

Each parsed artifact includes:
\begin{itemize}
    \item the sequence of actions extracted from the visible model answer;
    \item parser diagnostics such as missing actions, unknown action names, wrong arity(number of parameters of an action), markdown removal or compressed-action expansion;
    \item a separate reasoning-derived action sequence when provider-side reasoning text contains a candidate plan;
    \item a pointer to the corresponding raw artifact.
\end{itemize}

This layer bridges the gap between unstructured natural language and the action sequence that can be submitted to validation. If no usable action can be extracted, the benchmark records a parser failure and the case can still be scored without invoking VAL on an empty plan.

\subsubsection{Parsed File Structure}

Listing~\ref{lst:parsed_json_structure} shows the current parsed artifact schema. The \texttt{raw} subsection represents the official plan extracted from the visible answer; the \texttt{reasoning} subsection is diagnostic and is used to study whether a plan was present only in reasoning text.

\begin{lstlisting}[language=Python, caption={Structure of parsed JSON file}, label={lst:parsed_json_structure}]
{
  "model_id": #LLM model selected,
  "task_id": #problem and file selected,
  "protocol_id": #ierative repair or direct plan,
  "task_family": #PDDL domain problem selected,
  "tier": #easy, medium or hard depending on the instance,
  "instance_id": #instance selected,
  "generation_time_seconds": #sum of all the times of the single iterations,
  "raw_output_path": #path of the raw output ot the same instance,
  "attempts": [ #various iteration attempts
    {
      "iteration": #N. of current iteration,
      "generation_time_seconds": #time for the generation of the iteration,
      "parsed_plan": {
        "raw": {
          "actions": [ 
            #actions extracted from raw file
          ],
          "format_issues": [ #types of format issues if there are any ],
          "contains_reasoning": #true or false,
          "source_kind": #in this case is raw file
        },
        "reasoning": { #specific reasoning view
          "actions": [
            #reasoning actions extracted from raw file
          ],
          "format_issues": [ #format issues, some examples
            "unknown_action_names_found",
            "actions_embedded_in_text",
            "ambiguous_compressed_action"
          ],
          "source_ref": {
            "artifact": "raw",
            "field": "generation.reasoning_text"
          }
        }
      }
    },
  ]
}

\end{lstlisting}

\subsection{Scored}

The final layer, \texttt{outputs/scored/}, contains the normalized outcome of the case. If the parser extracts at least one action, the action sequence is validated with VAL. If parsing fails, the scored artifact stores a normalized \texttt{parse\_error} result instead. In both cases, the scored layer provides the fields used by the analysis notebooks and report metrics.

Each scored artifact includes:
\begin{itemize}
    \item the final \texttt{solved} decision for the benchmark instance;
    \item the number of iterations used and whether the iteration limit was reached;
    \item aggregate metrics such as \texttt{validity\_at\_1}, \texttt{validity\_at\_k}, plan length, repair success and failure type;
    \item one validation-style record for each attempt;
    \item repair feedback and reasoning-plan diagnostics when the iterative protocol or reasoning extraction uses them.
\end{itemize}

\subsubsection{Raw Text Validation}
After raw text is produced the validation is done in two ways: complete one shot validation and prefix validation where the one shot validation is done by validating the whole plan once the LLM finishes producing it and the prefix validation is conduced by validating the subplans by adding each a new action on top of the previous one.
This procedure is employed because we discovered that sometimes the LLMs output a plan that could result bigger than the actual plan needed.

\subsubsection{Scored File Structure}

Listing~\ref{lst:scored_json_structure} shows the main scored fields. The VAL-specific validation fields are nested under \texttt{validation\_result}; prefix-validation, reasoning, and feedback fields belong to the attempt record itself.

\begin{lstlisting}[language=Python, caption={structure of a scored JSON file.}, label={lst:scored_json_structure}]
{
  "model_id": #LLM selected ,
  "task_id": #domain and instance selected,
  "protocol_id": #iterative repair or direct plan,
  "solved": #true or false,
  "iterations_used": #number of iterations,
  "metrics": {
    "validity_at_1": #if valid with just the first iteration,
    "validity_at_k": #if valid at which itaeration,
    "iterations_to_valid": #n. of iterations to obtain a valid plan,
    "repair_success": #yes or no if iterative repair is on,
    "plan_length": #N. of actions in the plan,
    "error_type": #if error happens which one
    "hit_iteration_limit": #Current limit is 8, if reached without valid plan this becomes true
  },
  "raw_output_path": #path of the raw output ot the same instance
  "parsed_output_path": #path of the parsed output ot the same instance
  "attempts": [
    {
        "iteration": #current iteration,
        "generation_time_seconds": #time used for the generation of teh iteration,
        "validation_result": {
        "valid": #true or false,
        "status": #status of the validation,
        "error_type": #type of error, ex: unsatisfied_goal,
        "feedback_text": #feedbakc text containing also validator output,
        "failed_step": null,
        "failed_action": null,
        "goal_satisfied": #true or false,
        "plan_length": #N of plan's actions,
        "validation_time_ms": #time needed for validation,
        "raw_validator_output": #fro example: executed successfully,
        "details": {
            #details of the validator, for example external or internal validator
        },
        
      "first_valid_prefix_length": #lenght of the first subplan valid within the raw plan,
      "first_valid_plan_text": #text of the valid plan (obtained with first valid prefix) as a sequence of actions,
      "final_plan_valid": #true or false,
      "extra_actions_after_first_valid": #N of action done after valid subplan,
      
      "reasoning_validation_result": {
        "valid": #true or false,
        "status": #status of reasoning plan,
        "error_type": #error type of reasoning,
        "feedback_text": #feedback text for next iteration ,
        "failed_step": null,
        "failed_action": null,
        "goal_satisfied": #true or false,
        "plan_length": #N of reasoning plan's actions,
        "validation_time_ms": #time necessary for validation,
        "raw_validator_output": #path of the validation done by validator,
        "details": {
          #details of the validator, for example external or internal validator
        }

        "reasoning_first_valid_prefix_length": #lenght of the first subplan valid within the reasoning plan,
        "reasoning_first_valid_plan_text": ##text of the valid reasoning plan as a sequence of actions,
        "reasoning_final_plan_valid": #true or false,
        "reasoning_extra_actions_after_first_valid": null,
        "reasoning_candidate_count": #N of candidate plans parsed from reasoning,
        "reasoning_valid_candidate_count": #N of valid candidate plans parsed from reasoning,
        "reasoning_selected_candidate_index": #the index of the selected plan,
        "reasoning_selected_candidate_truncated": #true or false,
        "feedback_to_next_iteration": #feedback test to be used from next iteration
      },
    }
  ]
}
\end{lstlisting}
\newpage
