\section{Output Validation Using VAL}

After the benchmark has obtained a textual answer from an LLM and extracted the
candidate PDDL actions, the plan still has to be checked against the formal
domain and problem definition. This is the role of VAL in the framework. The
parser can say whether the model produced something that looks like an action
sequence; VAL decides whether that sequence is actually executable and whether
it reaches the goal.

\subsection{Role of VAL in the Benchmark}

VAL is used as an external symbolic validator. For each benchmark case, the
framework provides VAL with three inputs:

\begin{itemize}
    \item the PDDL domain file, containing predicates, actions, preconditions,
    and effects;
    \item the PDDL problem file, containing objects, initial state, and goal;
    \item the candidate plan extracted from the LLM output.
\end{itemize}

This separation is important because the benchmark does not rely on textual
judgement to decide whether a plan is correct. A model can produce a plausible
or well explained answer, but the answer is counted as solved only if the final
action sequence validates against the corresponding PDDL task.

\subsection{Validator Setup and Preflight}

VAL is not a Python dependency of the project. The benchmark can also run a task preflight before launching model generation.
In this stage the framework checks the selected domain and problem files with
VAL. This prevents malformed or unsupported PDDL inputs from being interpreted
later as model failures. If VAL cannot be used, or if the preflight fails,
the error is stored as an orchestration problem instead of being mixed with the
LLM planning results.

\subsection{Validation Flow}

The validation path is the same for every model backend. The backend changes how
the answer is generated, but not how the answer is checked:

\begin{center}
\texttt{LLM output}
$\rightarrow$ \texttt{parsed actions}
$\rightarrow$ \texttt{temporary plan file}
$\rightarrow$ \texttt{VAL(domain, problem, plan)}
$\rightarrow$ \texttt{ValidationResult}
$\rightarrow$ \texttt{scored artifact}
\end{center}

The temporary plan file is created only to invoke VAL with the expected command
line interface. After validation, the framework normalizes the validator output
into a common result schema. This makes the rest of the benchmark independent
from the exact textual format printed by the external validator.

\subsection{Normalized Validation Result}

The normalized validation record stores both the final decision and the
diagnostic information needed for analysis and repair.\\
A parse failure and a validation failure are not the same event. If no clean
PDDL action sequence can be extracted, VAL cannot judge the plan semantics and
the attempt is treated as a formatting or parsing failure. If actions are
extracted but VAL rejects them, the attempt reached the symbolic validation
stage and the error describes a planning failure or a validator-side technical
problem.

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Field} & \textbf{Meaning} \\
\midrule
\texttt{valid} & Boolean result used to decide whether the candidate plan solved the task. \\
\texttt{status} & Normalized status such as \texttt{valid}, \texttt{invalid}, \texttt{parse\_error}, \texttt{timeout}, or \texttt{validator\_error}. \\
\texttt{error\_type} & Compact failure class used by metrics and repair feedback. \\
\texttt{failed\_step} & First failing step reported by VAL, when available. \\
\texttt{failed\_action} & Problematic action extracted from the validator output, when available. \\
\texttt{goal\_satisfied} & Whether the final state satisfies the goal. \\
\texttt{plan\_length} & Number of action lines submitted to validation. \\
\texttt{validation\_time\_ms} & Time spent in the external validator call. \\
\texttt{raw\_validator\_output} & Original VAL output kept for debugging and later inspection. \\
\bottomrule
\end{tabularx}
\caption{Main fields stored in the normalized validation result.}
\label{tab:validation-result-fields}
\end{table}

\subsection{Prefix Validation}

The framework does not validate only the full action sequence. It validates
every non-empty prefix of the parsed final answer this provides a more precise
view of the model behaviour.\\
For example, a model may reach the goal after five actions and then continue
with unnecessary or invalid actions, in that case validating only the complete
sequence would mark the attempt as invalid, while prefix validation reveals that
a valid solution was present inside the answer. The scored artifact therefore
records the first valid prefix length, the first valid plan text, whether the
final full plan is valid and how many extra actions appear after the first
valid prefix.\\
This distinction is useful for the later reasoning analysis: it separates models
that never find a valid plan from models that find a plan but fail to stop cleanly
after satisfying the goal.

\subsection{Failure Types and Repair Feedback}

VAL failures are mapped into a small shared taxonomy. The most relevant classes
are:

\begin{itemize}
    \item \texttt{invalid\_precondition}: an action is not applicable in the
    current state;
    \item \texttt{unsatisfied\_goal}: the sequence executes but does not reach
    the required goal;
    \item \texttt{unknown\_action}: the plan contains an action not defined in
    the domain;
    \item \texttt{syntax\_error}: the plan format is not accepted by the
    validator;
    \item \texttt{timeout}, \texttt{validator\_unavailable}, and
    \texttt{validator\_crash}: technical validation failures, kept separate
    from planning mistakes.
\end{itemize}

In the iterative repair protocol, this normalized failure information is turned
into feedback for the next model attempt. The feedback can include the validation
status, the error type, the failed step, the failed action and a compact
description of the validator output; this makes the repair loop grounded in
symbolic validation rather than in a generic request to try again.

\subsection{Feedback Generation for Iterative Repair}

The repair feedback is generated as a deterministic diagnostic step over the
previous attempt, it is not a generic request to try again: the feedback selects
the most relevant failure signal and turns it into a constrained instruction for
the next model call.

\subsubsection{Inputs Used to Build Feedback}

The feedback builder can use the previous final answer, the previous reasoning
text when present, parser issues, the raw-plan validation result, the
reasoning-plan validation result, prefix-validation fields and alignment
signals between the final answer and the reasoning trace.\\
The protocol may also add a custom feedback prompt before the generated diagnostic text.

\subsubsection{Reasoning-Text Parsing as Feedback Input}

The \texttt{reasoning\_text} field is parsed through a separate diagnostic path.
This is a deliberate design choice: the visible final answer remains the only
official plan used to decide whether the benchmark case is solved, while the
reasoning trace is used to understand whether a useful plan was present before
the model wrote the final answer.\\
In other words, reasoning extraction can change the repair feedback, but it does not replace the raw final answer as the scored
submission.

The parsing starts from \texttt{generation.reasoning\_text} in the raw artifact.
The parser first removes Markdown fences, reads the domain action schemas and,
when available, the problem objects. It then extracts only domain-valid action
forms: the first token must be an action declared in the PDDL domain and the
number of arguments must match the action arity. If an argument is missing, the
parser fills it only when the problem contains exactly one compatible object for
the required type; this prevents the reasoning parser from inventing ambiguous
groundings.\\
Unknown action names, wrong arity, empty expressions, embedded prose and ambiguous compressed actions are kept as \texttt{format\_issues}.

Reasoning traces often contain more than one possible plan: an early attempt, analternative and then a final candidate, for this reason the parser does nottreat the entire reasoning text as one flat sequence but it builds candidate actionsegments, marks whether a segment appears near a final-answer marker, detectstruncated candidates, and may build a composite candidate when nearby action-likefragments look like parts of the same plan.\\
Composite construction stops at strong textual boundaries such as \texttt{alternative}, \texttt{maybe}, \texttt{instead}, \texttt{check}, \texttt{verify}, \texttt{wrong} or \texttt{invalid}. Safe shorthand forms are expanded only when they map to
unambiguous domain actions, for example local repeats such as \texttt{(action) x3}, \texttt{repeat 3 times (action)}, and one-argument compressed forms such as \texttt{b2 up x3}.

Listing~\ref{lst:reasoning_parser_code_refs} shows the main code references for this path: the extracted points to the parser entry point, the reasoning-specific section builder and the candidate extractor used for reasoning traces.

\begin{lstlisting}[language=Python, caption={Code references for reasoning-text parsing.}, label={lst:reasoning_parser_code_refs}]
# Benchmark_Framework/evaluators/parser.py

def parse_plan_text(raw_text, reasoning_text="", domain_text=None, problem_text=None):
    schemas = _parse_domain_actions(domain_text)
    objects_by_type = _parse_problem_objects(problem_text)
    return ParsedPlan(
        raw=_raw_section(raw_text, schemas, objects_by_type),
        reasoning=_reasoning_section(reasoning_text, schemas, objects_by_type),
    )

def _reasoning_section(reasoning_text, schemas, objects_by_type):
    if not reasoning_text.strip():
        return {"actions": [], "format_issues": ["reasoning_text_empty"]}

    cleaned_text, fence_issues = _strip_markdown_fences(reasoning_text)
    actions, extraction_issues = _extract_actions(
        cleaned_text,
        schemas,
        objects_by_type,
        select_candidate=True,
    )
    return {"actions": actions, "format_issues": fence_issues + extraction_issues}

def extract_reasoning_candidates(reasoning_text, domain_text=None, problem_text=None):
    schemas = _parse_domain_actions(domain_text)
    objects_by_type = _parse_problem_objects(problem_text)
    cleaned_text, fence_issues = _strip_markdown_fences(reasoning_text)
    candidates, shared_issues = _extract_action_candidates(
        cleaned_text,
        schemas,
        objects_by_type,
    )
    return candidates
\end{lstlisting}

After extraction, the runner validates the reasoning candidates with the same VAL
prefix-validation routine used for the final answer. Each candidate is checked as a complete sequence and through all non-empty prefixes then the selected diagnostic candidate is chosen by a deterministic priority order: prefer a fully valid candidate; otherwise prefer a candidate with a valid prefix; penalize truncated candidates; prefer candidates near final-answer markers; prefer candidates with fewer extra actions after the first valid prefix; then use source order, fewer
format issues and candidate length as tie breakers.\\
If the best reasoning candidate has a valid prefix but invalid trailing actions, the runner can shorten the diagnostic reasoning plan to that first valid prefix before storing it in the attempt record.

Listing~\ref{lst:reasoning_feedback_bridge_code_refs} shows where this parsed
reasoning signal becomes an input for the next feedback prompt. The key point is
that \texttt{reasoning\_validation\_result},
\texttt{reasoning\_first\_valid\_prefix\_length},
\texttt{reasoning\_final\_plan\_valid},
\texttt{reasoning\_candidate\_count} and
\texttt{reasoning\_selected\_candidate\_index} are stored in the same attempt
record that is later passed to \texttt{build\_repair\_feedback(...)}.

\begin{lstlisting}[language=Python, caption={Code references connecting reasoning parsing to repair feedback.}, label={lst:reasoning_feedback_bridge_code_refs}]
# Benchmark_Framework/runner/run_case.py

parsed_plan = _parse_generation_output(
    raw_text,
    generation.get("reasoning_text", ""),
    task_inputs.get("domain_text"),
    task_inputs.get("problem_text"),
)

reasoning_selection = _select_reasoning_candidate(
    validator,
    task_spec,
    _extract_reasoning_candidates(
        generation.get("reasoning_text", ""),
        task_inputs.get("domain_text"),
        task_inputs.get("problem_text"),
    ),
)

if reasoning_selection:
    selected_candidate = reasoning_selection["candidate"]
    parsed_plan["reasoning"]["actions"] = selected_candidate.get("actions", [])
    parsed_plan["reasoning"]["format_issues"] = selected_candidate.get("format_issues", [])

attempt_record["reasoning_validation_result"] = reasoning_summary["validation_result"]
attempt_record["reasoning_first_valid_prefix_length"] = reasoning_summary["first_valid_prefix_length"]
attempt_record["reasoning_final_plan_valid"] = reasoning_summary["final_plan_valid"]

feedback = build_repair_feedback(
    attempt_record=attempt_record,
    feedback_prompt=feedback_prompt,
    config=protocol_spec.repair_feedback,
)

# Benchmark_Framework/evaluators/repair_feedback.py

reasoning_valid = bool(
    attempt_record.get("reasoning_final_plan_valid")
    or reasoning_result.get("valid")
)

if reasoning_valid and not raw_valid:
    core = (
        "A valid action sequence was decoded from the reasoning text, "
        "but the final answer/raw plan did not validate. "
        "This points to a transfer problem from reasoning to final answer."
    )
\end{lstlisting}

This bridge explains how the final \texttt{feedback\_to\_next\_iteration} text is produced.\\
The feedback is not generated from the raw reasoning prose directly, it is generated from normalized diagnostics derived from that prose: the parsed reasoning actions, candidate-selection metadata, VAL validation results, prefix validation results and alignment between \texttt{parsed\_plan.raw.actions} and \texttt{parsed\_plan.reasoning.actions}. Therefore, when the reasoning trace contains a valid plan but the final answer does not, the feedback frames the error as a transfer problem. When reasoning and final answer contain mostly the same actions but VAL reports an \texttt{invalid\_precondition}, the feedback points to ordering or applicability. When reasoning and final answer diverge strongly, the feedback asks the model to rebuild one coherent plan and make the final answer match it.\\
If no reasoning plan is available, the repair message falls back to the raw parser issues and the validator result.

\subsubsection{Decision Order}

The feedback is written by following a priority order where technical failures are
handled first: if generation failed, VAL was unavailable, or validation timed
out, the feedback states that the attempt could not yet be judged as a confirmed
planning mistake and asks for a clean plan that can be validated again.

Formatting and parsing failures are handled next: if the final answer is empty,
does not contain a valid action sequence or mixes reasoning text into the final
answer the feedback focuses on output format rather than plan semantics: one
parenthesized PDDL action per line, with no surrounding prose.

If the answer reached semantic validation the feedback uses the normalized VAL fields. An \texttt{invalid\_precondition} error points to action applicability or ordering. An \texttt{unsatisfied\_goal} error points to a plan that executes but does not reach the goal.\\
If VAL exposes a failed step or failed action that information is included so the next attempt can repair the concrete failing part of the sequence.

After this the feedback checks whether prefix validation or reasoning validation changes the diagnosis. If a valid prefix exists but the final answer continues with extra actions the model is told to stop at the first valid solution. If the reasoning trace contains a valid plan but the final answer is invalid, the model is told to transfer the reasoning plan into the final answer. If reasoning and final answer are inconsistent, the model is asked to rebuild one coherent corrected sequence.


\subsubsection{Feedback Templates}

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Detected condition} & \textbf{Feedback content} \\
\midrule
Technical validation failure & Explain that the attempt could not be judged as a planning mistake and request a clean plan for re-validation. \\
Parser or format failure & Ask for only one parenthesized PDDL action per line, without explanations or markdown. \\
Reasoning valid, final answer invalid & Ask to copy the valid reasoning plan into the final answer. \\
Valid prefix with extra actions & Ask to stop immediately after the goal is reached. \\
\texttt{invalid\_precondition} & Ask to reorder or replace actions so every precondition is satisfied. \\
\texttt{unsatisfied\_goal} & Ask to extend or modify the plan until the goal condition is reached. \\
Low reasoning/final-answer alignment & Ask to rebuild one coherent plan and make the final answer match it. \\
\bottomrule
\end{tabularx}
\caption{How validation and reasoning diagnostics shape iterative-repair feedback.}
\label{tab:repair-feedback-flow}
\end{table}

\subsubsection{Iteration Budget and the Eight-Attempt Cap}

The repair loop also has an iteration budget. This budget is different from the prompt-token budget: it limits how many complete generate--parse--validate--repair cycles can be executed for a single benchmark instance: in the implemented iterative-repair protocol the maximum is eight attempts, as shown in Listing~\ref{lst:iterative_repair_iteration_budget}. The same cap is used for all models and domains so that retry-dependent success is comparable across runs.

\begin{lstlisting}[caption={Iteration budget in the iterative-repair protocol.}, label={lst:iterative_repair_iteration_budget}]
# Benchmark_Framework/protocols/iterative_repair.yaml

evaluation:
  max_iterations: 8
  require_final_plan_only: true
\end{lstlisting}

The choice of eight it is a bounded experimental compromise: Valmeekam et al.
(2023, Section 5.2 and Table 4) motivate verifier-assisted repeated
backprompting: an external checker can be used to return failure information and
measure whether later attempts improve. At the same time, Stechly, Marquez and
Kambhampati (2023, Section 4) show that, in iterative reasoning settings,
criticism or feedback can be largely irrelevant and later attempts can behave
like independent re-samples rather than genuine corrections.\\
For this reason the benchmark should allow enough attempts to observe possible feedback-driven correction, but it should not allow an unbounded retry search.

The analysis notebook uses this literature to justify a finite range of roughly 3 to 15 iterations: fewer than three attempts would mostly test one-shot performance and leave little room for repair, while values much larger than fifteen would mainly increase compute cost and inflate success through repeated sampling.\\
The value \texttt{max\_iterations = 8} is chosen as a conservative middle point in that range. It is high enough to observe early and medium-depth correction but low enough that a model cannot rely on unlimited sampling until it eventually emits a
valid plan.

This design is also tied to the metrics. First-attempt success rate
(\texttt{FASR}) isolates the retry-free planning signal.\\
Iteration-weighted success rate (\texttt{IWSR}) discounts late wins by assigning a success at attempt k a contribution of \(1/k\). Retry gap (\texttt{SR - FASR}) measures how much of the final success rate depends on retries.\\ 
Consequently, an eighth-attempt solution is still recorded as valid, but it is interpreted as much weaker evidence of planning competence than a first- or second-attempt solution. If no valid plan is found within the eight attempts, the scored artifact records that the run hit the iteration limit.

\subsubsection{Prompt-Budget Control and Final Constraint}
The previous raw answer and reasoning text can be included in the feedback
prompt, but they are truncated with separate limits. This keeps the useful
evidence from the failed attempt while avoiding a prompt dominated by old
outputs. After the diagnostic text is assembled, every branch ends with the same
final constraint: the next answer must contain only the complete PDDL action
sequence, one parenthesized action per line, with no explanation, headings,
markdown, numbering, or comments.

\subsection{Reasoning Validation}
Some models expose a separate reasoning trace in addition to the final answer.
The benchmark keeps the final answer as the official plan used for scoring, but
it can also validate an action sequence extracted from the reasoning text. This
second validation is diagnostic: it helps identify cases where the model derives
a valid plan in its reasoning but fails to transfer it correctly into the final
answer.

The scored result is therefore intentionally conservative: the official solved
flag comes from the final answer, not from a hidden or separate reasoning plan.
Reasoning validation is preserved to explain model behaviour, not to replace
the answer that would actually be submitted by the model.

This distinction avoids giving credit for hidden reasoning when the final answer
is invalid, while still preserving useful information for the reasoning
capability analysis in the following section.

\subsection{Metrics Derived from Validation}

The validation result is the source for the core execution metrics. A case is
considered solved only when the normalized validation result is valid. From this
record the framework derives \texttt{validity\_at\_1},
\texttt{validity\_at\_k}, \texttt{repair\_success},
\texttt{iterations\_to\_valid}, \texttt{plan\_length},
\texttt{error\_type}, and whether the run hit the iteration limit.

Therefore, VAL is not only a final correctness check. It is the component that
connects raw model generations to comparable benchmark scores.

\subsection{Limitations}

The validation layer deliberately checks formal plan correctness, not the
quality of the natural-language reasoning. It also depends on the availability
and behaviour of an external executable, so timeout and validator errors are
reported separately from genuine planning failures. Finally, the parser that
normalizes VAL output uses conservative textual heuristics; for this reason the
raw validator output is stored together with the normalized fields whenever
manual inspection is needed.

