\section{Conclusion}
To conclude, it is worth highlighting the successful creation of a scalable, expandable and modular benchmark for evaluating the PDDL planning capabilities of Large Language Models. The framework was designed to be extensible to a wide range of problems and models, with a strong emphasis on producing easily interpretable and traceable results.

\subsection{General Conclusions on LLM Planning Capabilities}
Our evaluation of several models highlighted critical insights about the current state of LLM-based planning: the main finding is that while LLMs often handle PDDL syntax and action vocabulary, their ability to perform robust, long-horizon, and goal-directed reasoning is still fragile.
Models frequently generate action sequences that are syntactically correct and locally executable, as pointed out by generally high \texttt{Executability Ratios} and low \texttt{Hallucination Rates}, but these plans often fail to achieve the global goal, leading to low overall \texttt{Success Rates} due to a non-coherent world model. The significant gap between first-attempt success (\texttt{FASR}) and the final success rate (\texttt{SR}) would suggest a strong dependency on iterative repair, but by looking at \texttt{IWSR} we can see that current models often behave more like efficient-correctors than deterministic planners.

\subsection{Limitations}
It is also worth noticing the limitations of this study: the evaluation is constrained by the available infrastructure, since these LLM models require significant resources to run effectively.
The observed performance is also tied to the specific prompts and iterative repair strategies employed: different approaches could yield different results.\\
It is also important to dive into the fact that the way we can interact with LLMs is less user friendly than the way we can interact with popular conversational AIs like Claude or chatGPT. This is due to a design nature that causes an interaction limitation because the LLMs don't have memory between iterations, this is the direct cause of the need to implement the entire feedback system.

\subsection{Hints for Future Work}
The modularity of the benchmark creates several possibilities for future work and improvement:
\begin{itemize}
    \item \textbf{Single Domain Testing:} The set of PDDL problems could be reduced to a single reasoning type and the difficulty increased gradually to see where a given LLM starts to fail at planning.
    
    \item \textbf{Advanced Feedback Prompting:} Future work could explore more sophisticated strategy for feedback generation. This includes implementing more informative feedback mechanisms for the iterative repair loop, providing models with specific details about precondition failures (e.g., "action `(A)` failed because precondition `(P)` was false") to better assess their self-correction abilities. A possible way that could be explored when dealing with reasoning mmanagment, and therefore feedback, is to implement a LLM to extract more effectively the plan from the reasoning instead of using an hard coded approach.
    
    \item \textbf{Broader Model Evaluation:} The benchmark could be used to evaluate a broader and wider set of models, to build a more comprehensive landscape of LLM planning skills.

    \item \textbf{Single Model Multi-Release Evaluation:} The benchmark could be used to evaluate different releases of the same model, to test how its planning capabilities evolve across versions.
    
    \item \textbf{Deeper Analytical Insights:} The current exhaustive amount of data gathered by the outputs could be used also to perform deeper correlational studies, which could reveal fundamental strengths and weaknesses of different architectures.

\end{itemize}
