Within-domain ranking and its direction rules, the per-domain heatmap normalisation caveat, rank variance (generalist vs specialist), Spearman domain correlation, the failure-mode taxonomy quadrants, and the seven literature-grounded capability profiles with their NaN-aware assignment logic.

A single overall leaderboard hides the most scientifically interesting fact about a model:\emph{where} it is good. This view documents the three machinery pieces that turn per-domain metrics into comparative claims --- within-domain ranking, rank variance, and the Spearman domain-correlation matrix --- and then the failure-mode taxonomy and the seven literature-grounded capability profiles that convert those numbers into a named diagnosis.

\subsubsection{Within-domain ranking: why rank, not raw value}

Absolute metric values are not comparable across domains --- a 0.30 success rate in a hard transport domain and a 0.30 in an easy gripper domain mean different things. The framework therefore ranks models \textbf{inside each domain separately}, on every metric in \code{RANK\_METRICS}, then works with ranks. Ranking removes the domain-difficulty offset and answers the only question that survives it:\emph{among the models that faced this exact domain, who placed where?}

The direction of "good" is metric-specific and is encoded once in \code{RANK\_METRICS}. Maximise for Success\_Rate, FASR, IWSR, Exec, PAS, CoT\_Alignment; minimise for Hallucination, Retry\_Gap, Temporal\_Distance. The rank call makes the direction explicit:

\begin{lstlisting}
for metric, direction in RANK_METRICS.items():
    rank_table[metric] = sub[metric].rank(
        ascending = direction == "min",   # min-metrics: smallest value = rank 1
        method    = "min",                # ties share the best rank
        na_option = "bottom",             # missing values rank last, never first
    )
# rank 1 = best in this domain on this metric
\end{lstlisting}

\begin{bmkeybox}{The rationale of the procedure}
Ranking is a deliberate refusal to average raw scores across heterogeneous domains. "Model A beats Model B" is only ever asserted

\emph{within a domain on a named metric}

; a global statement is then built by aggregating those local verdicts (mean rank, rank variance), never by averaging raw metric values that live on incomparable scales.

\code{na\_option="bottom"}

is a fairness rule: a model that produced no usable plan cannot be handed a good rank by accident.
\end{bmkeybox}

\subsubsection{The ranking heatmap and its hard caveat}

The \code{domain\_ranking\_heatmap} renders one sub-grid per domain (rows = models, columns = RANK\_METRICS). To share one colour scale across domains with different model counts, each raw rank is min-max normalised inside its domain to [0,1]:

\begin{bmformulabox}
\begin{align*}
\text{norm\_rank}
&= \frac{\text{rank} - 1}{\max(\text{n\_models} - 1, 1)} \\
&\in [0,1],
\quad 0 = \text{best in domain},
\quad 1 = \text{worst in domain} \\
&\text{Colormap: RdYlGn\_r, green = best.}
\end{align*}
\end{bmformulabox}

\begin{bmcavbox}{Do not read normalised cells as cross-domain quality}
The normalisation is

\textbf{per-domain}

. A 0.00 (green) cell means "best

\emph{among the models present in this domain}

", not "good in absolute terms". A model can be deep green in a domain where every model failed. Compare cells

\emph{down a column within one domain block}

; never compare the same model's colour across two different domain blocks as if it were a score. The colour encodes local standing, not capability.
\end{bmcavbox}

\subsubsection{Rank variance: generalist vs specialist}

Rank variance compresses "where is this model good" into one number. For each model it takes the standard deviation of its within-domain rank across domains, per metric, then averages those across metrics (\code{mean\_rank\_std}). A model that places \textasciitilde{} 3rd everywhere has near-zero variance (a generalist); a model that wins one domain and trails in another has high variance (a domain specialist).

\begin{lstlisting}
var_df = rank_within_domain.groupby("Model")[RANK_METRICS_cols].std()
var_df["mean_rank_std"] = var_df[RANK_METRICS_cols].mean(axis=1)
var_df = var_df.sort_values("mean_rank_std", ascending=False)  # specialists on top
\end{lstlisting}

phi\_4 carries the highest \code{mean\_rank\_std \(\approx\) 2.394}(most domain-specific --- strong in a domain or two, weak elsewhere) while gpt\_oss sits lowest at \code{\(\approx\) 1.349}(most consistent across domains). High variance is not "bad": it flags a model whose single overall number is least trustworthy, because its behaviour is domain-contingent.

\begin{bmcavbox}{Spec says "normalised rank", code uses raw rank}
The plot docstring describes rank variance as the std of

\emph{normalised}

rank, but the implementation computes std over the

\emph{raw}

integer ranks in

\code{rank\_within\_domain}

(the normalisation is applied only inside the heatmap). With a fixed model count the two differ by a constant factor, so the ordering of models is unaffected --- but the absolute

\code{mean\_rank\_std}

values are on the raw-rank scale, not [0,1]. Worth aligning if you ever compare variance magnitudes across runs with different model counts.
\end{bmcavbox}

\subsubsection{Domain correlation: redundancy vs complementarity}

The \code{domain\_correlation} plot is a model\(\times\)model Spearman \(\rho\) matrix built from each model's vector of per-domain success rates. Two models with \(\rho\)\(\approx\)1 succeed and fail on the same domains --- they are \emph{redundant} probes of capability. Two with low or negative \(\rho\) stress different things --- they are \emph{complementary}, and a benchmark wants both. It is a tool for curating the model set and the domain set, not for ranking.

\begin{lstlisting}
sr_pivot = df_metrics.pivot_table(index="Model", columns="Domain", values="Success_Rate")
corr = sr_pivot.T.corr(method="spearman")   # Model \(\times\) Model rank correlation
\end{lstlisting}

\begin{bmcavbox}{Correlation needs a domain spread}
Spearman over success-rate vectors is only meaningful with several domains and some variance in them. In the bundled run the domain count is small and many models sit at SR\(\approx\)0, so \(\rho\) is dominated by ties and should be read as illustrative of the method, not as a stable finding.
\end{bmcavbox}

\subsubsection{Breadth of competence: stacked PS and the SR heatmap}

\code{ps\_by\_domain\_stacked} stacks each model's per-domain Planning Score contributions into one bar. Two models can reach the same total bar height by opposite routes: a tall single segment (high PS concentrated in one domain) versus many even segments (competence spread across domains). The stack makes "broad vs concentrated" visible at a glance --- exactly the distinction the overall PS scalar destroys.

The \code{success\_rate\_heatmap} is the raw companion: model\(\times\)domain SR with no normalisation, so here --- unlike the rank heatmap --- cells \emph{are} comparable across the whole grid. Use it to spot empty columns (domains no model solved \(\rightarrow\) an infrastructure or difficulty problem, not a model problem) and empty rows (models that solved nothing).

\subsubsection{The failure-mode taxonomy plot}

This is the framework's single most diagnostic figure. It places every model on two orthogonal failure axes ---\textbf{hallucination rate}(x, grounding failure) against \textbf{PAS}(y, precondition awareness) --- encoding \textbf{FASR as bubble size} and \textbf{SR as colour}. The point is that "the model failed" is not one phenomenon: the two axes separate four qualitatively different failure regimes.

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
Quadrant & Reading & Diagnosis\\
\midrule
low halluc \(\cdot\) high PAS & grounded, legal names, but mis-orders actions & Sequencing failure (e.g. gemma)\\
\addlinespace
low halluc \(\cdot\) low PAS & legal names, no coherent state model & Vocabulary-only\\
\addlinespace
high halluc \(\cdot\) high PAS & invents actions yet shows some precondition sense & Confused / mixed\\
\addlinespace
high halluc \(\cdot\) low PAS & not reading the domain; everything ungrounded & No grounding\\
\bottomrule
\end{tabularx}
\end{table}

\begin{bmkeybox}{What the taxonomy lets you conclude}
Two models with identical success rates can sit in different quadrants --- and that difference is the actual research finding. In the bundled run gemma is the only sequencing-heavy model (failure breakdown 20 sequencing vs 28 fabrication, seq-proportion 0.417), so it lives in the grounded/high-PAS corner while the flat-zero models collapse toward the ungrounded corner. The taxonomy turns an undifferentiated pile of failures into named, citable failure phases.
\end{bmkeybox}

\subsubsection{The seven capability profiles}

The profiles are the taxonomy made formal: each is a conjunction of threshold conditions over \code{\{SR, FASR, IWSR, RG, IHR, ER, PAS, CoT, p\_valid\_k\_flat\}}, and each maps to a specific claim in the planning literature. They are the bridge from "here are nine numbers" to "this model is doing X".

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
Profile & Conditions & Interpretation \(\cdot\) reference\\
\midrule
\textbf{Genuine Planner} & SR>0.70 \(\cdot\) FASR>0.50 \(\cdot\) RG<0.10 \(\cdot\) IHR>0.90 \(\cdot\) CoT>0.70 & Reads the schema, solves first try, reasoning tracks the plan.\\ Strongest claim\\
\addlinespace
\textbf{Efficient Corrector} & SR>0.50 \(\cdot\) FASR<0.30 \(\cdot\) |SR-IWSR|\(\le\)0.10 \(\cdot\) 0.10\(\le\)RG\(\le\)0.30 & Rarely first-try, but repairs land early so IWSR stays near SR.\\ Stechly 2023 (verify via P(Valid|k))\\
\addlinespace
\textbf{Stochastic Searcher} & SR\(\ge\)0.30 \(\cdot\) FASR<0.20 \(\cdot\) RG>0.30 \(\cdot\) [opt] P(Valid|k) flat & Success inflated by retries; attempts behave like resampling.\\ Stechly 2023\\
\addlinespace
\textbf{Lucky Retriever} & SR>0.60 \(\cdot\) FASR>0.50 \(\cdot\) IHR<0.30 \(\cdot\) CoT<0.40 & Solves from surface patterns; weak grounding, post-hoc reasoning.\\ Kambhampati 2024 \(\cdot\) Valmeekam obfuscation\\
\addlinespace
\textbf{Understander} & SR<0.20 \(\cdot\) FASR<0.10 \(\cdot\) ER>0.70 \(\cdot\) IHR>0.90 \(\cdot\) PAS>0.70 & Knows vocabulary \&local preconditions, fails long-horizon order.\\ Valmeekam Phase 2 \(\cdot\) Dziri 2023\\
\addlinespace
\textbf{Vocabulary-Only} & SR<0.15 \(\cdot\) IHR>0.90 \(\cdot\) 0.30\(\le\)ER\(\le\)0.70 \(\cdot\) PAS<0.30 & Legal action names, no coherent state-transition model.\\ Kambhampati 2024 \(\cdot\) Vafa 2024\\
\addlinespace
\textbf{No Grounding} & SR\(\le\)0.05 \(\cdot\) IHR<0.60 \(\cdot\) ER<0.30 \(\cdot\) PAS<0.30 & Not reading the formal domain at all; everything ungrounded.\\ Valmeekam 2023 \(\cdot\) McCoy 2024\\
\bottomrule
\end{tabularx}
\end{table}

\subsubsection{How a model is assigned a profile}

Assignment is deliberately conservative. Every profile is evaluated; the \textbf{first exact match}(all required conditions met, none unavailable) wins. If nothing matches exactly the model is labelled \code{Mixed/Unclassified} and reported against its \emph{closest} profile --- the one with the highest matched-condition count, ties broken by fewest required conditions.

\begin{lstlisting}
evaluations = [evaluate_profile_conditions(metrics, p) for p in PROFILE_DEFINITIONS]
exact = next((e for e in evaluations if e["exact"]), None)
if exact is None:
    best = max(evaluations, key=lambda e: (e["match_count"], -e["required_count"]))
    assigned = "Mixed/Unclassified"      # honest about partial matches
else:
    assigned = exact["profile"]
\end{lstlisting}

\begin{bmkeybox}{NaN-aware conditions: "unavailable" \(\neq\) "missing"}
A condition whose metric is NaN (e.g. CoT when reasoning was never produced) is marked

\emph{unavailable}

, not

\emph{failed}

. This is what stops a model being thrown out of "Genuine Planner" merely because CoT could not be computed --- and it is why the report exposes the full matched/missing/unavailable breakdown per profile, so you can see exactly how close a model sat to each boundary rather than trusting a bare label.
\end{bmkeybox}

\begin{bmcavbox}{Profiles are thresholds, not ground truth}
The cut-points (SR

>

0.70, IHR

>

0.90, ...) encode the literature's qualitative claims as hard numbers; they are defensible but not sacred. A model just under a boundary is reported as Mixed against its closest profile precisely so the threshold choice stays visible and auditable. Treat the label as a hypothesis the condition breakdown lets you check, not a verdict.
\end{bmcavbox}

\subsubsection{Reading the metric set together: the diagnostic framework}

No single metric is a verdict. The scientific meaning lives in \emph{combinations}--- and the within-domain ranking heatmap is where they are read side by side. There are three reading directions, and then a small set of metric pairs that must never be read in isolation.

\paragraph{Three ways to read the ranking heatmap.}

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
Direction & What it tells you\\
\midrule
A \textbf{row}(one model) & Scan all metrics: uniformly green = leads on every dimension here; uniformly red = worst on everything; a mixed row is a capability profile --- strengths and weaknesses at once.\\
\addlinespace
A \textbf{column}(one metric) & Which model dominates that metric. A column all one colour means the metric doesn't discriminate in this domain.\\
\addlinespace
\textbf{Cell blocks} & Green SR+FASR but red halluc+CoT = succeeding despite poor grounding (pattern-matching). Red SR but green exec+PAS = failing at the plan level yet showing process-level understanding.\\
\bottomrule
\end{tabularx}
\end{table}

\paragraph{The metric pairs to always read together.}

\textbf{Capability triangle --- SR + FASR + IWSR + RG}(is performance genuine or scaffolded by retriesc). Read SR and FASR first, then RG = SR - FASR, then IWSR position.

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
SR & FASR & RG & IWSR & Diagnosis\\
\midrule
High & High & \(\approx\)0--0.10 & \(\approx\)FASR & Genuine zero-shot planner --- retries irrelevant.\\
\addlinespace
High & Medium & 0.10--0.25 & \(\approx\)SR & Efficient corrector --- converges on attempts 2--3.\\
\addlinespace
High & Low & >0.30 & \(\approx\)FASR & Stochastic search --- SR inflated; late retries are re-samples.\\
\addlinespace
Medium & Low & 0.20--0.40 & mid & Partial capability, retry-scaffolded.\\
\addlinespace
Low & Low & \(\approx\)0--0.05 & \(\approx\)FASR & No capability --- nothing for retries to inflate.\\
\bottomrule
\end{tabularx}
\end{table}

\textbf{Failure anatomy --- ER + PAS}(most useful when SR<0.15: how deep does it run, and why did it stopc)

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
ER & PAS & What the model is doing & Valmeekam phase\\
\midrule
>0.70 & >0.70 & Runs deep, ordering errors near the goal --- understands domain, mis-sequences late. & Phase 2 (long-horizon)\\
\addlinespace
>0.70 & <0.30 & Runs deep but fabricates states --- knows vocabulary, invents transitions. & Phase 1/2 boundary\\
\addlinespace
<0.30 & >0.70 & Fails early, but on ordering of the few real steps. Rare. & Phase 2 + vocab issues\\
\addlinespace
<0.30 & <0.30 & Fails immediately, invents states from step 1. & Phase 1 (vocabulary)\\
\addlinespace
Medium & NaN & Mixed; PAS undefined --- check halluc gate and temporal distance. & Mixed\\
\bottomrule
\end{tabularx}
\end{table}

\textbf{Reasoning test --- SR + CoT} and \textbf{refinement pair --- IHR + PAS}(genuine vs retrieved success; and which kind of failure).

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
Pair & Signature & Meaning\\
\midrule
SR high \(\cdot\) CoT>0.70 & --- & Reasoning guides grounded plans --- strongest evidence of planning (Lanham).\\
\addlinespace
SR high \(\cdot\) CoT<0.40 & --- & Fake reasoning --- correct plans, post-hoc CoT (Kambhampati).\\
\addlinespace
SR low \(\cdot\) CoT>0.70 & --- & Faithful but failing --- reasoning grounded, execution fails.\\
\addlinespace
IHR>0.90 \(\cdot\) PAS>0.70 & top-left (good) & Knows vocabulary, ordering errors only --- most fixable.\\
\addlinespace
IHR>0.90 \(\cdot\) PAS<0.30 & bottom-left & Knows words, lacks world model.\\
\addlinespace
IHR<0.60 \(\cdot\) PAS low & bottom-right (worst) & Near-random output --- no grounding, no causal model.\\
\bottomrule
\end{tabularx}
\end{table}

\begin{bmcavbox}{IHR is a gate, not just a metric}
If inverse-hallucination rate falls below \textasciitilde{} 0.6, ER, PAS and CoTaL are all unreliable: the model is failing at vocabulary, so the downstream causal metrics have nothing trustworthy to measure. Always clear the IHR gate before interpreting the process metrics.
\end{bmcavbox}

\paragraph{The diagnostic tree.}

The pairs above compose into a single ordered procedure --- read the metrics in this sequence, taking the branch that matches, and the model classifies itself.

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
IHR<0.60 & STOP. Not reading the schema --- vocabulary failure. Do not interpret exec/PAS/CoT.\\
\addlinespace
IHR 0.60--0.90 & Proceed with caution; note partial vocabulary issues.\\
\addlinespace
IHR>0.90 & Proceed to Level 1.\\
\bottomrule
\end{tabularx}
\end{table}

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
SR<0.15 & Almost no valid plans --- skip to Level 4; the question is which failure dominates.\\
\addlinespace
SR 0.15--0.50 & Partial capability --- Level 2.\\
\addlinespace
SR>0.50 & Can plan here --- Level 2 to test authenticity.\\
\bottomrule
\end{tabularx}
\end{table}

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
RG<0.10 & Retry-independent (FASR\(\approx\)SR) --- Level 3 + Level 5.\\
\addlinespace
RG 0.10--0.30 & Moderate dependency --- report SR and FASR; Level 3.\\
\addlinespace
RG>0.30 & Strong dependency --- SR inflated, use FASR; confirm via flat P(Valid|k).\\
\bottomrule
\end{tabularx}
\end{table}

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
IWSR \(\approx\) FASR & All success on attempt 1 --- confirm with declining P(Valid|k).\\
\addlinespace
IWSR near SR & Efficient corrector (attempts 2--3) --- verify with rising P(Valid|k).\\
\addlinespace
IWSR near FASR, high RG & Late-retry luck (attempts 4--5) --- stochastic; flat P(Valid|k).\\
\bottomrule
\end{tabularx}
\end{table}

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
ER>0.70 \(\cdot\) PAS>0.70 & Long-horizon sequencing failure --- Phase 2 (Valmeekam).\\
\addlinespace
ER>0.70 \(\cdot\) PAS<0.30 & Runs deep but fabricates --- world-model incoherence (Vafa).\\
\addlinespace
ER<0.30 \(\cdot\) PAS<0.30 & Fails immediately --- Phase 1 fabrication.\\
\addlinespace
ER 0.30--0.70 & Mixed --- check temporal distance (low = near-miss, high = broken causal model).\\
\bottomrule
\end{tabularx}
\end{table}

\begin{table}[h]\small
\centering
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash} X >{\raggedright\arraybackslash} X@{}}
\toprule
CoT>0.70 \(\cdot\) SR high & Genuine deliberation --- strongest result.\\
\addlinespace
CoT<0.40 \(\cdot\) SR high & Fake reasoning --- post-hoc CoT (Kambhampati); report as caveat.\\
\addlinespace
CoT>0.70 \(\cdot\) SR low & Faithful but failing --- likely long-horizon (if ER also high).\\
\addlinespace
CoT<0.40 \(\cdot\) SR low & Complete grounding failure --- return to Level 0 for a masked vocab issue.\\
\bottomrule
\end{tabularx}
\end{table}
